MSF_Groupwork-NBA_shotlog

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
additional explanation for defender proximity

##Defender Proximity Analysis  
###Closest Defender Distance vs Shot Accuracy
####Introduction  
This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome. The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring. This is because the shooter is practically in "free" condition without any physical interference.

@@

#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
  scale_x_discrete("Closest Defender Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)) +
  ylab("Shooting Accuracy (%)") +
  ggtitle("Closest Defender Distance vs Shooting Accuracy")
```
###Plot Result
  The explanation from the plot above are as follows:   
  1. If the defender stands really close to the shooter (0 to 2 feet), the accuracy varied between **42%** to **50%**  
  2. If the defender distance range from 2 (two) to 4 (four) feet, the accuracy rate for this distance range from **44%** to **49%**  
  3. If the defender distance range from 4 (four) to 6 (six) feet, the accuracy rate for this distance range from **41%** to **46%**
  4. If the defender distance range from 6 (six) to 8 (eight) feet, the accuracy rate for this distance range from **37%** to **42%**
  5. If the defender distance range from 8 (eight) to 10 (ten) feet, the accuracy rate for this distance range from **37%** to **54%**
  6. For defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
  7. For the distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)

###Findings  
  As the analysis shown, the accuracy rate of every player in the last season was somehow **unaffected** by the defender distance. For defender distance starting from 0 (zero) to For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance). It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result. This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy.

###Limitations  
  As the dataset did not hold many data points for more than 20 feet defender distance, the accuracy analysis for these distances are limited. The analysis could be enhanced further by including additional season data. That way, there will be more data point for further distance and possibly would produce stronger pattern as the short distance shown.
```{r}
#Calculate accuracy against defender distance
#Group the defender distance level

distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
head(distance_level)

#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  scale_y_continuous(labels = percent) +
  scale_x_discrete("Closest Defender Distance", breaks=c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, Inf)) +
  ylab("Shooting Accuracy (%)") +
  ggtitle("Closest Defender Distance vs Shooting Accuracy")

#Calculate accuracy against shooting distance
#Group the shooting distance level
shooting_distance_level <- as.data.frame(table(factor(ShotDistance)))
head(shooting_distance_level)

>>>>

##Defender Proximity Analysis  
###Closest Defender Distance vs Shot Accuracy
####Introduction  
This analysis wants to calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome. The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring. This is because the shooter is practically in "free" condition without any physical interference.

@@

#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
  scale_x_discrete("Closest Defender Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)) +
  ylab("Shooting Accuracy (%)") +
  ggtitle("Closest Defender Distance vs Shooting Accuracy")

```
###Plot Result
  The explanation from the plot above are as follows:   
  1. If the defender stands really close to the shooter (0 to 2 feet), the accuracy varied between **42%** to **50%**  
  2. If the defender distance range from 2 (two) to 4 (four) feet, the accuracy rate for this distance range from **44%** to **49%**  
  3. If the defender distance range from 4 (four) to 6 (six) feet, the accuracy rate for this distance range from **41%** to **46%**
  4. If the defender distance range from 6 (six) to 8 (eight) feet, the accuracy rate for this distance range from **37%** to **42%**
  5. If the defender distance range from 8 (eight) to 10 (ten) feet, the accuracy rate for this distance range from **37%** to **54%**
  6. For defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
  7. For the distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)

  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:
```{r}
distance_level$Var1 <- as.numeric(distance_level$Var1)
distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1

distanceZerotoAll <- c()
accuracyDist1 <- c()
for (i in 0:nrow(distance_level)){
  distanceZerotoAll <- c(distanceZerotoAll, distance_level[i, "Var1"])
  accuracyDist1 <- c(accuracyDist1, distance_level[i, "Accuracy"])
}

dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
cor.test(dfZerotoFive$distanceZerotoAll, dfZerotoFive$accuracyDist1)
```
###Correlation Test Result  
  The correlation test result produced slightly high number of correlation between defender distance with shooting accuracy for all shot distance attempt in the last seasons's NBA (0.59). If the findings made only based on this number only, there would be obvious that further defender distance will **automatically** increase shooting accuracy. However, after the analysis went into smaller piece of distance, ther result will be difference. For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:

```{r}
distanceZerotoEight <- c()
accuracyDist2 <- c()
for (i in 0:80){
  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
}

dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
```
  Therefore,

###Discussion  
  As the analysis shown, the accuracy rate of every player in the last season was somehow **affected** by the defender distance for certain range. For defender distance starting from 0 (zero) to 8 (eight) feet For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance). It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result. This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy.

###Limitations  
  As the dataset did not hold many data points for more than 20 feet defender distance, the accuracy analysis for these distances are limited. The analysis could be enhanced further by including additional season data. That way, there will be more data point for further distance and possibly would produce stronger pattern as the short distance shown.

###Shooting Distance vs Shot Accuracy
####Introduction  
This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome. The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring. This is because the shooter is practically in "free" condition without any physical interference.

###Methods  
The steps to calculate this measurement are as follows:  
  1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
  2. Select distinct of **all** closest defender distance  
  3. Count the occurence fOr every distance  
  4. Calculate how many succesfull shot scored from every distance point  
  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
  6. Plot the result
```{r}
#Calculate accuracy against shooting distance
#Group the shooting distance level
shooting_distance_level <- as.data.frame(table(factor(ShotDistance)))
head(shooting_distance_level)
